---
title: "R Notebook"
output:
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Importing gun violence 


```{r}
tx <- read_csv("data-raw/gun_accidents_tx.csv") %>% clean_names()
```

## Cleaning

### Columns additions

Adds an id so we can export parts and merge the back to the correct records. Sets city as value of city_or_county, which is cleaned in the next step. Adds county which is filled later.

```{r}
tx <- tx %>% 
  mutate(id = row_number(),
         county = "",
         city = city_or_county
         )
```

## Clean city

```{r}
tx <- tx %>% 
  mutate(
    city = str_replace(city, ".*county.*", "")
  )
```

## Clean county

There are six records that have a county listed in `city_or_county`. We populate the `county` column only when the word phrase " (county)" is in the `city_or_county` field. Then we remove the the phrase " (county)". Note the double-escape for parenthesis.

```{r}
tx <- tx %>% 
  mutate(
    county = case_when(
      grepl("county", city_or_county) ~ city_or_county
    )
  ) %>%
  mutate(
    county = str_remove(county, " \\(county\\)")
  )
```

### Clean dates

```{r}
tx <- tx %>% 
  mutate(
    incident_date = mdy(incident_date)
  )
```


### Reorganize the columns

Peek at the names

```{r}
tx %>% names()
```

Reselect to change the order

```{r}
tx <- tx %>% 
  select(
    id,
    incident_date,
    address,
    city,
    county,
    state,
    number_killed,
    number_injured,
    operations,
    city_or_county
  )
```

## Export the file

```{r}
tx %>% write_rds("data-out/01_tx.rds")
tx %>% write_csv("data-out/01_tx.csv")
```

