---
title: "A look at geocodio"
output:
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

There is a file `data-raw/01_tx_gun_accidents_geocodio.csv` that is the result of sending `data-out/01-tx.csv` through [geocodio](https://www.geocod.io/). I added the census tract, demographics and economic columns. It brought back a ton of info ... more than necessary. Will try to cut that down here.

They also have an R package [rgeocodio](https://github.com/hrbrmstr/rgeocodio) that is worth exploring. Might be able to specify just the columns needed.


```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(janitor)
```

## Import the geocodio-returned data

```{r}
tx_geocodio <- read_csv("data-raw/01_tx_gun_accidents_geocodio.csv") %>% clean_names()
```

It looks like this failed to parse 4 row: 14, 124, 139, 175. They were short columns?

## Peek at data

Man, there are a lot of columns

```{r}
tx_geocodio %>% head()
```


## Select the columns we need

I'm going to print the names out to csv to take a closer look at them.

```{r}
tx_geocodio %>% names() %>% as_tibble() %>% write_csv("data-out/02_tx_geocodio_names.csv")
```

```{r}
tx_geocodio_select <- tx_geocodio %>% 
  select(
    id,
    incident_date,
    address,
    city,
    county,
    state,
    number_killed,
    number_injured,
    operations,
    city_or_county,
    number,
    street,
    city_2,
    zip,
    county_2,
    acs_economics_number_of_households_total_value,
    acs_economics_number_of_households_total_margin_of_error,
    acs_economics_median_household_income_total_value,
    acs_economics_median_household_income_total_margin_of_error,
    accuracy_score,
    accuracy_type,
    state_2,
    latitude,
    longitude,
    state_fips,
    county_fips,
    place_fips,
    census_tract_code,
    census_block_code,
    census_block_group,
    full_fips
  ) %>% 
  rename(
    households_total = acs_economics_number_of_households_total_value,
    households_moe = acs_economics_number_of_households_total_margin_of_error,
    median_income_total = acs_economics_median_household_income_total_value,
    median_income_moe = acs_economics_median_household_income_total_margin_of_error
  )
```


## Export the geocoded file

```{r}
tx_geocodio_select %>% 
  write_csv("data-out/02_tx_geocodio_select.csv")

tx_geocodio_select %>% 
  write_rds("data-out/02_tx_geocodio_select.rds")
```


## Understanding the accuracy of geocoding results

Cribbed from [geocodio docs](https://www.geocod.io/docs/#accuracy-score).

Consider `accuracy_type` first as it has a heirarchy, with the best being rooftop and worst being state. 


| Value                 | Description                                                                                         |
|-----------------------|-----------------------------------------------------------------------------------------------------|
| rooftop               | The exact point was found with rooftop level accuracy                                               |
| point                 | The exact point was found from address range interpolation where the range contained a single point |
| range_interpolation   | The point was found by performing address range interpolation                                       |
| nearest_rooftop_match | The exact house number was not found, so a close, neighboring house number was used instead         |
| street_center         | The result is a geocoded street centroid                                                            |
| place                 | The point is a city/town/place                                                                      |
| state                 | The point is a state                                                                                |

Each geocoded result is returned with an accuracy score, which is a decimal number ranging from 0.00 to 1.00. The higher the score, the better the result.

Generally, accuracy scores that are larger than or equal to 0.8 are the most accurate, whereas results with lower accuracy scores might be very rough matches.

That said, a 1.0 perfect match for the "state" type is not a good result, because that is just the center of the state.

If you want to consider which of the results have accurate data to the zip code level, I would exclude some records in a couple of ways:

- Exclude all that have `accuracy_type` of "place" or "state". That leaves 137 results.
- Then consider the `accuracy_score`. 0.5 or greater gives you 133 records. Go to 0.8 and it drops to 96. Could always hand-check those lower-rated results to see if they past muster.

```{r}
'%ni%' <- Negate('%in%')
tx_geocodio_select %>% 
  filter(accuracy_type  %ni% c("place", "state"),
         accuracy_score >= .8)
```

So, in the end the geogoding results are not any better than with MapChi ... 
